Forecasting the Changing Dynamics of US Energy Product Exports to OECD Markets: The Role of the Russian-Ukraine Conflict and Future Scenarios

### 2023-05-18 draft.Rmd
updated

---
title: "FM project"
output:
  pdf_document: default
  toc: yes
  toc_depth: 2
  theme: simplex
  date: "2023-04-25"
  fig_width: 7
  fig_height: 6  
  fig_caption: true
  html_document: null

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

@@

## Analysing U.S. energy exports

## 1.Introduction

Energy commodities play a pivotal role in the global economy, as they are essential for economic growth and development. They are required for transportation, manufacturing, heating, cooling, and many other critical activities that drive economic activity.

@@

In this study, we will analyze the impact of the war in Ukraine on US energy exports to OECD Europe. Specifically, we will focus on three main energy products: crude oil, petroleum products, and natural gas. By examining trends in these exports, we can better understand how the geopolitical events in Ukraine have affected the energy market in Europe. Additionally, we will investigate changes in OECD Europe's dependence on Russian gas to see how energy dynamics in the region have shifted.

## 2. Time Series Graphics

@@

#### 2.1 Dataset

@@

```{r, data_prep, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
library(fpp3)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(scales)

@@

#### 2.2 Time Series Plots

@@

#### 2.3 Including Plots

@@

```{r, time series overview, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}

@@

## 3. Time series decomposition

#### 3.1. Adjustments

@@

#### 3.2. Transformations

@@

#### 3.3 Decomposition

@@

## 4. The forecaster’s toolbox

#### 4.1. Finding the fit

@@

#### 4.2. Residual diagnostics

@@

```{r setup, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}

@@

fit %>% select(additive_ETS) %>% report()
additive_ets<-fit %>% select(additive_ETS)
addi_fc<- additive_ets %>% forecast(h=36)
accuracy(additi)
addi_fc %>% 
  autoplot(
    crude_oil_bx) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))
```

@@

For the training dataset, the additive model had the lowest ME (Mean Error) of -0.036 while RMSE (Root Mean Squared Error), MAE (Mean Absolute Error) and most of the other metrics are relatively similar when comparing Additive and Random models which are two best performing ETS models..

Additionally, the additive model had a lower ACF1 when comparing to other two models.A lower ACF1 (Autocorrelation Function at Lag 1) indicates a better model fit.

Considering these metrics, the additive model demonstrated superior performance in terms of forecasting accuracy for the crude oil time series. Hence, we selected the additive model (ETS=AAA) as the most appropriate ETS model for our analysis.

@@

```{r applying, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
#ETS model selection by R
crude_fit <- crude_oil_bx %>%
  model(auto_crude_ets = ETS(Total_Exports))
#checking the accuracy

accuracy(crude_fit)
report(crude_fit)
report(additive_ets)

auto_crude_fc<-crude_fit %>% select(auto_crude_ets) %>% forecast(h=36)

auto_crude_fc %>% 
  autoplot(crude_oil_bx) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))

@@

After applying the ETS() function to choose an appropriate model for the crude oil time series, we obtained the selected ETS model as <ETS(A,A,N)>. To assess the accuracy of the selected model, we calculated various metrics using the accuracy() function. The results indicate that the chosen model performs reasonably well in terms of forecasting accuracy. The ME (Mean Error) is close to zero, indicating that, on average, the model's forecasts are unbiased. The RMSE (Root Mean Squared Error) is 8.7, which represents the average magnitude of the forecast errors. The MAE (Mean Absolute Error) is 6.64, which measures the average absolute difference between the model's forecasts and the actual values. Finally, the ACF1 (Autocorrelation of Forecast Errors) is 0.129, representing the correlation between successive forecast errors. Overall, the selected ETS model demonstrates satisfactory accuracy, aligning with the main features observed in the crude oil time series.

@@

In contrast, the auto_crude_ets model (ETS(A,A,N)) chosen through information criteria has different parameters and characteristics, including smoothing parameters and initial states. Its information criteria values are lower than those of the additive model.

@@

5.3 Checking the residuals

```{r bash, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
#Plotting Residual
gg_tsresiduals(additive_ets)
gg_tsresiduals(crude_fit)
```

In context of white noise, in the case of the additive model and crude_fit, the significant values at lag 12 and lag 15 respectively suggest that there is some correlation or pattern present in the residuals at those specific lags.

@@

5.4 Comparison of ETS model to the simple models

```{r Comparion of Simple Models with ETS Model, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}

combined_accuracy <- accuracy(crude_oil_fit) %>%
  arrange(.model) %>%
  select(.model, .type, RMSE, MAE, MAPE, MASE, RMSSE) %>%
  bind_rows(
    accuracy(additive_ets),
    accuracy(crude_fit)
  )

combined_accuracy

Among the forecast models evaluated, the additive_ETS and auto_crude_ets models demonstrate significantly better performance compared to the Drift, Mean, Naive, and Seasonal Naive models. The additive_ets model exhibits lower values for RMSE, MAE, MAPE, and MASE, indicating higher accuracy in predicting future values. Additionally, the model has a lower RMSSE value, suggesting a better fit to the data. While the ME is close to zero, indicating a minor bias in the forecasts, the MPE is negative, implying a slight underestimation of actual values on average. Also, low ACF1 suggests a weak positive correlation in the residuals at Lag 1. Similarly, the auto_crude_ets model shows favorable results with comparable accuracy metrics. Overall, the additive_ets and auto_crude_ets models outperform the other models, making them more suitable choices for forecasting.

@@

Yet to be started

>>>>

---
title: "Forecasting the Changing Dynamics of US Energy Product Exports to OECD Markets: The Role of the Russian-Ukraine Conflict and Future Scenarios"
date: "2023-04-25"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    fig_width: 7
    fig_height: 6  
    fig_caption: true
  fontsize: 11pt
  geometry: margin=1in

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

\newpage

@@

THINK ABOUT PROPHET FORECASTING MODEL
\newpage

## 1. Introduction

\linebreak  
Energy commodities play a pivotal role in the global economy, as they are essential for economic growth and development. They are required for transportation, manufacturing, heating, cooling, and many other critical activities that drive economic activity.

@@

In this study, we will analyze the impact of the war in Ukraine on US energy exports to OECD Europe. Specifically, we will focus on three main energy products: crude oil, petroleum products, and natural gas. By examining trends in these exports, we can better understand how the geopolitical events in Ukraine have affected the energy market in Europe. Additionally, we will investigate changes in OECD Europe's dependence on Russian gas to see how energy dynamics in the region have shifted.   
\linebreak  
\linebreak     
## 2. Time Series Graphics

@@

### 2.1 Dataset

@@

```{r, data_prep, echo=FALSE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
library(fpp3)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(tidyr)
library(lubridate)
library(scales)

@@

### 2.2 Time Series Plots

@@

### 2.3 Including Plots

@@

```{r,Times series overivew, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}

@@

## 3. Time series decomposition   
### 3.1. Adjustments

@@

### 3.2. Transformations

@@

### 3.3 Decomposition

@@

## 4. The forecaster’s toolbox   
### 4.1. Finding the fit

@@

### 4.2. Residual diagnostics

@@

```{r setup ETS, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}

@@

addi_fc<- fit %>% select(additive_ETS) %>% forecast(h=36)

addi_fc %>% 
  autoplot(
    crude_oil_bx) +
  labs(
    y = "Thousand Barrels",
    title = "Forecasts for U.S. Crude oil exports to OECD Europe") +
  guides(colour = guide_legend(title = "Forecast"))
```

@@

For the training dataset, the additive model had a lower RMSE (Root Mean Squared Error) of 4.51 compared to 5.12 for the multiplicative model and 4.78 for the random model. The MAE (Mean Absolute Error) was also lower for the additive model (3.52) compared to the multiplicative model (3.99) and random model (3.65).

Additionally, the additive model had a lower MAPE (Mean Absolute Percentage Error) of 7.37% compared to 8.32% for the multiplicative model. The MASE (Mean Absolute Scaled Error) value for the additive model (0.415) was lower than that of the multiplicative model (0.470), indicating better forecasting accuracy.

Considering these metrics, the additive model demonstrated superior performance in terms of forecasting accuracy for the natural gas time series. Hence, we selected the additive model (ETS=AAA) as the most appropriate ETS model for our analysis.

@@

```{r applying, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}

@@

After applying the ETS() function to choose an appropriate model for the crude oil time series, we obtained the selected ETS model as <ETS(A,Ad,N)>. To assess the accuracy of the selected model, we calculated various metrics using the accuracy() function. The results indicate that the chosen model performs reasonably well in terms of forecasting accuracy. The ME (Mean Error) is close to zero, indicating that, on average, the model's forecasts are unbiased. The RMSE (Root Mean Squared Error) is 4.70, which represents the average magnitude of the forecast errors. The MAE (Mean Absolute Error) is 3.64, which measures the average absolute difference between the model's forecasts and the actual values. The MPE (Mean Percentage Error) is -1.33%, indicating a slight underestimation of the crude oil exports. The MAPE (Mean Absolute Percentage Error) is 7.87%, providing a measure of the relative forecasting accuracy. The MASE (Mean Absolute Scaled Error) is 0.430, suggesting that the model's forecasts outperform a naive seasonal method. The RMSSE (Root Mean Squared Scaled Error) is 0.464, which compares the model's performance to that of a naïve random walk model. Finally, the ACF1 (Autocorrelation of Forecast Errors) is 0.183, representing the correlation between successive forecast errors. Overall, the selected ETS model demonstrates satisfactory accuracy, aligning with the main features observed in the crude oil time series.

@@

In contrast, the auto_crude_ets model (ETS(A,Ad,N)) chosen through information criteria has different parameters and characteristics, including smoothing parameters and initial states. Its information criteria values are lower than those of the additive model.

@@

5.3 Checking the resuduals

```{r bash, echo=TRUE, results='markup',warning=FALSE,message=FALSE, fig.align='center', background= 'gray'}
#Plotting Residual
```

In context of white noise, in the case of the additive model and crude_fit, the significant values at lag 9 and lag 17 respectively suggest that there is some correlation or pattern present in the residuals at those specific lags.

@@

@@
